An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-01-07 Articles/Article_EnerEcon_2014/EconGrowth3.Rnw
Added notes about precision.

<<CES_sigma_delta_graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models.">>=
standardScatterPlot(data=subset(cesCoeffsData, nestStrParen != "(iK + iL) + ()"),
                    mapping=aes(delta, sigma), facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

@@

>>>>

<<CES_sigma_delta_graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models.">>=
standardScatterPlot(data=filter(cesCoeffsData, nestStrParen != "(iK + iL) + ()"),
                    mapping=aes(delta, sigma), facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

@@

**** From the above discussion, 
it is clear that when $\min{sse}$ occurs at any boundary 
(with the exception of $\sigma, \, \sigma_1 = 0$, $\rho, \, \rho_1 = -1$), 
it will be difficult to estimate many CES parameters with precision.
It is concerning that so many of our CES results end up on boundaries. 
I don't think we know why we hit the boundaries so frequently. 
I think it's real, i.e., lowest $sse$ occurs at the boundary. 
But, I can't know for sure, because the \texttt{cesEst} function
is a black box to us.
One thing in our favor is that we do a thorough grid search in $\rho$ and $\rho_1$ first.
Presumably, the grid search finds any potential 
$sse$ valleys before starting the full gradient search. 
This is probably an issue to be discussed with Randy. --MKH ****

**** We do a thorough job of fitting along the boundaries for the Cobb-Douglas function.
Perhaps we should do something similar for the CES model? 
I.e., we could fit the CES model unconstrained.
We could set up a list of models to try along each boundary. 
Similar to Cobb-Douglas, if an unconstrained CES point estimate lands beyond the boundary,
we could simply choose the lowest $sse$ anywhere on the boundary as the winner. --MKH ****

**** This raises another question,
namely whether we should be reporting coefficients that are, by definition unknowable.
In the worst case, when $\delta = 0$ 
(as it is for the fit to historical data for 
US;
(le)k nesting, Equation~\ref{eq:CESlek}; in Figure~\ref{fig:CES_sigma_delta_graph}), 
CES simplifies to $y = \gamma A l$.
All of $\sigma$, $\delta$, $\sigma_1$, and $\delta_1$
are eliminated from the model
and are, therefore, unknowable.  
At present, we report the values of uknowable coefficients on the resample graphs.
One approach would be to decide \emph{not} to report unknowable values.
But, there are a few resample models that are rather close to but not exactly at $\delta = 1$
in Figure~\ref{fig:CES_sigma_delta_graph}.
Should we report parameters for the resample models where $\delta \ne 1$? --MKH ****
